---
title: "LA-EDA"
author: "Jatin-Dheeraj"
date: "2023-05-14"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

## we using readr package to read a CSV file called "housing.csv" from your local file system and then viewing the data using the View() function.

```{r echo=TRUE, warning=FALSE}
library(readr)
housing <- read_csv("C:/Users/HP/Downloads/archive (2)/housing.csv")
View(housing)
```

------------------------------------------------------------------------

## The str() function in R provides a concise summary of the structure of a given object.

```{r echo=TRUE}
str(housing)


```

## table(housing\$households) in R creates a frequency table that shows the number of occurrences of each unique value in the households column of the housing data frame.

```{r echo=TRUE}
options(max.print = 30)
table(housing $ households)
```

## By changing the options("max.print") value, you can set the default number of rows to display.

```{r echo=TRUE}
options(max.print = 30)


```

## class(housing)" will return the class of the object "housing".

```{r echo=TRUE}
class(housing)
```

## head(housing)" will display the first few rows of the "housing" dataset or data frame, allowing the user to quickly preview the structure and contents of the data

```{r echo=TRUE}
head(housing)

```

## summary(): Provides a summary of a data frame

```{r echo=TRUE}
summary(housing)
```

## dim(): Shows the dimensions of a data frame

```{r echo=TRUE}
  dim(housing)
```

## nrow(): Shows the number of rows in a data frame

```{r echo=TRUE}
  nrow(housing)
```

## ncol(): Shows the number of columns in a data frame

```{r echo=TRUE}
  ncol(housing)

```

## names(): Shows the column names of a data frame

```{r echo=TRUE}
names(housing)
```

## tail(): Shows the last few rows of a data frame

```{r echo=TRUE}
  tail(housing)
```

## subset(): Extracts a subset of a data frame based on a condition

```{r echo=TRUE}
  subset(housing)
```

## select(): Selects specific columns from a data frame

> ## Select columns

```{r echo=TRUE}

library('dplyr')
housing %>% select(2,3)
```

## Select columns by list of index or position

```{r echo=TRUE}
housing %>% select(c(2,3))
```

## Select columns by index range

```{r echo=TRUE}
housing %>% select(2:3)
```

## Select columns by label'total_rooms','housing_median_age'

```{r echo=TRUE}
 housing %>% select('total_rooms','housing_median_age')
```

## Select columns except 'total_rooms','housing_median_age'

```{r echo=TRUE}
housing %>% select(-c('total_rooms','housing_median_age'))

```

## Select columns between 'total_rooms','housing_median_age'

```{r echo=TRUE}
housing %>% select('total_rooms':'housing_median_age')
```

## Select columns starts with a string

```{r echo=TRUE}
housing %>% select(starts_with('l'))
```

## Select columns that ends with a string

```{r echo=TRUE}
housing %>% select(ends_with('e'))
```

## Select columns that contains

```{r echo=TRUE}
housing  %>% select(contains('p'))
```

## Select all numeric columns

```{r echo=TRUE}
housing %>% select_if(is.numeric)
```

## subset(): Extracts a subset of a data frame based on a condition

# Syntax of subset()

subset(x, subset, select, drop = FALSE, ...)

## subset by row name

```{r echo=TRUE}
subset(housing, subset=rownames(housing) == '41') 
```

## subset row by of row values

```{r echo=TRUE}
subset(housing, rownames(housing) %in% c('41','52','42'))
```

## subset() by Condition or Column Value

```{r echo=TRUE}
  subset(housing, latitude=='37.83')
```

## subset by multiple conditions

```{r echo=TRUE}
  subset(housing, population =='754' | total_bedrooms == '311')
```

## subset columns

```{r echo=TRUE}
  subset(housing,housing_median_age=='51',select=c('longitude','latitude'))
```

## subset columns by index

```{r echo=TRUE}
  subset(housing,housing_median_age=='51',select=c(1,2,3))
```

## filter(): Filters a data frame based on a condition

# filter() by row name

```{r echo=TRUE}

library('dplyr')
filter(housing, rownames(housing) == '51')

```

## filter() by column Value

```{r echo=TRUE}
  filter(housing, total_rooms == '896')
```

## filter() by list of values

```{r echo=TRUE}
  filter(housing, housing_median_age %in% c("21", "39", "33"))
```

## filter() by row number

```{r echo=TRUE}
  slice(housing, 5)
```

## arrange(): Sorts a data frame by one or more columns

# Syntax of arrange()

arrange(.data, ..., .by_group = FALSE)

## Using arrange in descending order

```{r echo=TRUE}
  temp <- housing %>% arrange(desc(latitude))
   temp
```

## Using arrange in descending order

```{r echo=TRUE}
  temp <- housing %>% arrange((housing_median_age))
  temp
```

## Using arrange by multiple columns

```{r echo=TRUE}
  temp <- housing %>% arrange(housing_median_age,desc(total_rooms))
   temp
```

## group_by(): Groups a data frame by one or more columns

# Syntax group by

group_by(.data, ..., add = FALSE)

## Load dplyr

```{r echo=TRUE}
  library(dplyr)
```

## group_by() on department

```{r echo=TRUE}
grp_tbl <- housing %>% group_by(housing_median_age)
grp_tbl
  
```

## summarise on groupped data.

```{r echo=TRUE}
  agg_tbl <- grp_tbl %>% summarise(sum(housing_median_age))
  agg_tbl
```

## Convert tibble to DataFrame

```{r echo=TRUE}
  df2 <- agg_tbl %>% as.data.frame()
  class(df2)
```

## Assign column Name to the aggregated column

## Group by on multiple column

```{r echo=TRUE}
  agg_tbl <- housing %>% group_by(housing_median_age) %>% 
               summarise(total_salary=sum(total_rooms))
agg_tbl
```

## cbind(): Concatenates two data frames by adding columns

## 

```{r echo=TRUE}
y1 <- c(9, 8, 7, 6, 5)    
  data_new1 <- cbind(housing, y1)     # cbind vector to dataframe
data_new1                             # Print data to RStudio console
```

## is.na(): Checks for missing values in a data frame

```{r echo=TRUE}
  is.na(housing)
```

## complete.cases(): Removes rows with missing values

```{r echo=TRUE}
  complete.cases(housing)
  
```

## mean(): Calculates the mean of a numeric vector

```{r echo=TRUE}
  mean(housing$median_income)
```

## median(): Calculates the median of a numeric vector

```{r echo=TRUE}
  median(housing$latitude)
```

## sd(): Calculates the standard deviation of a numeric vector

```{r echo=TRUE}
  sd(housing$housing_median_age)
```

## var(): Calculates the variance of a numeric vector

```{r echo=TRUE}
  var(housing$longitude)
```

## max(): Finds the maximum value in a numeric vector

```{r echo=TRUE}
  max(housing$total_rooms)
```

## min(): Finds the minimum value in a numeric vector

```{r echo=TRUE}
  min(housing$population)
```

## quantile(): Calculates the quantiles of a numeric vector

```{r echo=TRUE}
  quantile(housing$longitude)
```

## unique(): Finds the unique values of a vector or column in a data frame

```{r echo=TRUE}
  unique(housing$total_rooms)
```

## duplicated(): Identifies duplicated values in a vector or column in a data frame

```{r echo=TRUE}
  duplicated(housing$households)
```

## as.numeric(): Converts a factor or character vector or column in a data frame to a numeric vector

```{r echo=TRUE}
    # creating a character vector
x = c('1', '2', '3', '4', '5')

# implementing the as.numeric() function 
y = as.numeric(x) 

# printing the numeric vector
y

# checking if its a numeric matrix
is.numeric(y)
```

## To divide a vector into groups, use the split() function.

```{r echo=TRUE}
    split(housing, f = housing$ocean_proximity)
```

## Split data frame by ocean_proximity

```{r echo=TRUE}
    s <- split(housing, housing$ocean_proximity)
    s
```

## Add Column Using cbin() Function

```{r echo=TRUE}
percentage = c(76,86)
df2 <- cbind(housing, percentage)
df2

```

## Add Column in R using [] Notation

```{r echo=TRUE}
chapters = c(1,2,3,4,5,6,7,8,9,10)
df2['new'] <- chapters
df2

```

## \# Using \$ notation

```{r echo=TRUE}
  chapters = c("A","B","C")
df2$chapters <- chapters
df2

```

## length() function: returns the number of elements in a vector

```{r echo=TRUE}
length(housing)

```

## Filtering data based on conditions:

```{r echo=TRUE}
filtered_data <- subset(housing, housing_median_age > 20 & median_income >50000)
filtered_data
```

## Grouping data and calculating aggregate functions:

```{r echo=TRUE}
library(dplyr)
grouped_data <- housing %>% group_by(ocean_proximity) %>% summarise(avg_median_income = mean(median_income), max_median_house_value = max(median_house_value))
grouped_data
```

## Adding a new column to the dataset:

```{r echo=TRUE}
housing$new_column <- housing$total_rooms + housing $total_bedrooms
head(housing)

```

## Saving the modified dataset to a new CSV file:

```{r echo=TRUE}
write.csv(housing, file = "modified_dataset.csv", row.names = FALSE)

```

## Calculate the correlation between variables:

```{r echo=TRUE}
correlation_matrix <- cor(housing[, c("housing_median_age", "total_rooms", "total_bedrooms", "population", "households", "median_income", "median_house_value")])
correlation_matrix 

```

## Counting the number of unique values in a column:

```{r echo=TRUE}
unique_count <- length(unique(housing$ocean_proximity))
unique_count

```

## Rename a column:

```{r echo=TRUE}
names(housing)[names(housing) == "ocean_proximity"] <- "proximity_to_ocean"
head(housing)
```

## Create a new variable based on conditions:

```{r echo=TRUE}
housing$high_income <- ifelse(housing$median_income > 100000, "High", "Low")
head(housing)
```

## Compute a cross-tabulation of two variables:

```{r echo=TRUE}
cross_tab <- table(housing$latitude, housing$longitude)
head(cross_tab)
```

## Calculate the cumulative sum of a variable:

```{r echo=TRUE}
housing$cumulative_rooms <- cumsum(housing$total_rooms)
head(housing)

```

## Calculate the quantiles of a variable:

```{r echo=TRUE}
quantiles <- quantile(housing$median_house_value, probs = c(0.25, 0.5, 0.75))
quantiles

```

## Subset data based on multiple conditions:

```{r echo=TRUE}
subset_data <- housing[housing$housing_median_age > 30 & housing$population < 1000, ]
head(subset_data)
```

## If you want to randomly sample a specific number of rows from the dataset, you can use the sample_n() function from the dplyr package.

```{r echo=TRUE}
library(dplyr)
sample_n(housing, size = 10)


```

## Calculate the mode using the mode function from the pracma

```{r echo=TRUE}
library('pracma')

mode_value <- mode(housing$median_income)
mode_value
```

## Calculate the mode using the table function:

```{r echo=TRUE}
mode_value <- names(table(housing$median_income))[which.max(table(housing$median_income))]
mode_value
```

## Perform hierarchical clustering on numerical variables:

```{r echo=TRUE}
num_data <- housing[, c("housing_median_age", "total_rooms", "total_bedrooms", "population", "households", "median_income", "median_house_value")]
hclust_result <- hclust(dist(num_data))

hclust_result

```

## Calculate the Euclidean distance between two variables:

```{r echo=TRUE}
euclidean_distance <- dist(head(housing[, c("total_rooms", "total_bedrooms")]), method = "euclidean")
euclidean_distance 

```

## Calculate the percentage of missing values in each column:

```{r echo=TRUE}
missing_percentage <- colMeans(is.na(housing)) * 100
missing_percentage 
```

## Calculate the covariance between two variables:

```{r echo=TRUE}

covariance <- cov(housing$total_rooms, housing$median_house_value)
covariance
```

## Convert a variable to a different data type:

```{r echo=TRUE}
temp <- housing
temp$housing_median_age <- as.factor(temp$housing_median_age)
class(temp$housing_median_age)
```

## Adding a new column to the dataset:

```{r echo=TRUE}
temp$new_column <- temp$total_rooms + temp$total_bedrooms
head(temp)

```

## 

```{r echo=TRUE}

median_by_group <- housing %>%
  group_by(proximity_to_ocean) %>%
  summarise(median_median_income = median(median_income))
median_by_group

```

## Create a scatter plot matrix:

```{r echo=TRUE}
library(ggplot2)

scatter_plot_matrix <- ggplot(housing, aes(x = longitude, y = latitude)) +
  geom_point() +
  facet_grid(. ~ proximity_to_ocean) +
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Scatter Plot Matrix")

print(scatter_plot_matrix)

```

## Create a bar plot with error bars:

```{r echo=TRUE}


line_plot <- ggplot(housing, aes(x = housing_median_age, y = median_house_value, color = proximity_to_ocean)) +
  geom_line() +
  xlab("Housing Median Age") +
  ylab("Median House Value") +
  ggtitle("Line Plot")

print(line_plot)


```

## Create a scatter plot with a trend line:

```{r echo=TRUE}
library(ggplot2)
temp <- head(housing)
scatter_plot <- ggplot(temp, aes(x = total_rooms, y = median_house_value)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Total Rooms") +
  ylab("Median House Value") +
  ggtitle("Scatter Plot with Trend Line")

print(scatter_plot)

```

## Create a bar plot with stacked bars:

```{r echo=TRUE}




stacked_bar_plot <- ggplot(temp, aes(x = proximity_to_ocean, fill = housing_median_age)) +
  geom_bar(position = "stack") +
  xlab("Proximity to Ocean") +
  ylab("Count") +
  ggtitle("Stacked Bar Plot")

print(stacked_bar_plot)



```

## Create a pie chart:

```{r echo=TRUE}


pie_chart <- ggplot(temp, aes(x = "", fill = proximity_to_ocean)) +
  geom_bar(width = 1) +
  coord_polar("y", start = 0) +
  xlab("") +
  ylab("") +
  ggtitle("Pie Chart")

print(pie_chart)


```

## Create a scatter plot with jittered data points:

```{r echo=TRUE}


scatter_plot <- ggplot(temp, aes(x = total_rooms, y = median_house_value)) +
  geom_jitter() +
  xlab("Total Rooms") +
  ylab("Median House Value") +
  ggtitle("Scatter Plot with Jittered Data Points")

print(scatter_plot)


```
